---
title: "eWHORM -- Statistical considerations for setting up the simulation study"
author: MUW Team
format: 
  pdf:
    documentclass: scrartcl  
    toc: true
      # toc-depth: 3
      # toc-title: Contents
editor: visual
---

```{r, echo=F, warning=F, message=F}
library(lme4)
```

\newpage

## Preamble

Brief report to summarise the key points of the planned design and discuss the simulation study.

## Aims and scope

Preliminary study on the design of a randomized, placebo-controlled, double-blind, parallel-group phase-II study to investigate the efficacy and safety of oxfendazole (OXF) in subjects with onchocerciasis, loiasis, mansonellosis and/or soil-transmitted helminth (STH).

In particular, the goal of this report is to explore the state-of-the-art and establish the basis for a simulation study aiming at investigating  the  key operating characteristics of the
trial.

The simulations will optimize different design elements, including interim analyses and stopping rules, dose
selection at interim, (adaptive) randomization and treatment allocation, sample size reassessment and
subgroup analyses. Furthermore, the simulations will help in guaranteeing a prespecified statistical power, type 1 error rate control and ensuring the trial
integrity and validity. Adaptive decision rules will be pre-specified and interim and final analysis
will feed into the design. With specific statistical testing and estimation procedures accounting for adaptations such
as sample size reassessment and dropping of treatment arms, the type 1 error rate will be controlled and the sample
size reassessed. 

## Main elements of the design

### Sub-studies

Focusing on filarial and STH infections, we consider a basket trial to test OXF against four diseases (onchocerciasis, loiasis, mansonellosis and STH). 
The basket trial then considers four sub-studies corresponding to each of the previous-mentioned diseases. 

#### Endpoints

The efficacy endpoints of the  sub-studies are the following:  


-   OXF in Onchocerciasis patients
    - The efficacy endpoint will be the presence (or absence) of live female adult worms with normal embryogenesis  at month 12.

-   OXF in Loiasis patients:
    - The efficacy endpoint will be the presence (or presence) of L. loa mf at 12 months.
    <!-- after treatment. -->

-   OXF in mansonellosis patients:
    - The efficacy endpoint will be the presence (or absence) of mf at month 12.
    
-   OXF in T. trichiura patients:
    - The efficacy endpoint is the T. trichiura cure rate assessed at 14 - 21 days post-treatment using the duplicate Kato-Katz technique. Other endpoints are egg reduction rates (ERRs) and infection status and intensity of infection.
    
Notes: 1) For some of the diseases it will be important to perform subgroup analyses (for now not included here). 2) For mansonellosis, we will need to account for co-infections. In particular, mansonellosis impacts co-infections with M. tuberculosis and HIV.    
    
#### Treatments

In the three sub-studies with filarial patients (onchocerciasis, loiasis and mansonellosis), patients will receive either 400 mg OXF single dose (SD), 800 mg OXF SD, 800 mg OXF once daily for 5 days, or placebo. In the sub-study with trichuriasis patients, OXF will be administered at
400mg and 800mg SD.

- onchocerciasis, loiasis and mansonellosis: three treatments and placebo
- trichuriasis patients: two treatments and placebo

### Trial stages

The trial includes two-stages. 

- Stage 1. In the first stage, 120 patients per indication will be included. After stage 1 ends, an interim analysis will be performed in all sub-studies. According to that, the trial may be adapted (treatment arms can be stopped for futility or efficacy, sample size will be reassessed). 
- Stage 2. Trial continues to the second stage where up to 80 additional
patients may be allocated to the selected doses and the placebo group. 


Study duration will be up to 12 months, depending on the indication. 

### Trial objectives

The main objectives of the trial are the following:

- Primary objective: To demonstrate a positive dose-response relationship of OXF
in the primary endpoint for each of the four indications: onchocerciasis, loiasis, mansonellosis, trichuriasis/STH
infection. 

- Secondary objectives: To demonstrate the efficacy of specific doses. 

Further points to be considered include (not all included for now): the safety and
tolerability of OXF in helminth infected individuals, geographical region, PK parameters of OXF and their
relationship to efficacy response and safety and tolerability will be investigated.


## Design considerations in previous and/or published studies

Summary clinical papers


- Welsche et al. (2022)
  - The primary outcome was egg reduction rate (ERR) of T trichiura 14–21 days after treatment in the available case population.
  - The estimated sample size, based on simulations, resulted in 160 participants would be needed per group to ensure a power of 90% that the upper limit of the two-sided 95% CIs would exclude a difference of more than 2 percentage points (the non-inferiority margin) in favour of the ivermectin and albendazole combination
group, assuming a true ERR of 98% in both groups.
  - Analysis:
    - ERRs were calculated using geometric mean egg counts
    - To estimate the difference between ERRs and 95% CIs, a bootstrap resampling method with 5000 replicates was used.
    - Logistic regression with adjustment for baseline infection intensity, age, sex, and weight was also performed. For each follow-up timepoint, ERRs and cure rates were calculated.
    
- Hürlimann, et al. (2022)
  - The primary outcome was cure rate against T trichiura, defined as the proportion of participants with no eggs in their faeces 14–21 days after treatment, assessed by Kato-Katz thick smears
  - Secondary outcomes were the ERR against T trichiura; cure rates
and ERRs against A lumbricoides, hookworm, and S stercoralis; and infection status assessed by qPCR
  - Sample size based on the unpooled difference in proportions sample size formula. 143 participants
per treatment group
  - Analysis:
    - A melded binomial test with mid-p correction was used
to calculate differences in cure rates between the two
treatment groups and to estimate the corresponding
confidence intervals.
    - Geometric mean and arithmetic mean egg counts, and bootstrap as before. 


- Gardon, et al. (2002)
  - The primary outcome was the vital status of female
worms as assessed from the proportion and mean
number of female worms in the MD category per nodule.
  - For an expected difference of 30% in the proportion of
moribund and dead female worms between the four groups at the end of the trial, and for a patient drop-out
rate of 50%, 544 patients were needed, taking
into account the data in the published work for the
usual proportion of dead worms found before
treatment
  - Analysis:
    - calculated the mean numbers of microfilariae per
mg of skin with William’s geometric mean
    - tested for any difference between all the four groups, with
the chi2 test for proportions, and with ANOVA or the
Kruskal-Wallis rank-sum test for means.  Random-effects logistic-regression models   to
compare the effects of the three other treatment regimens
with that of the control group.

- Preprint paper pending to be included:
  - Zoleko-Manego et al  on loiasis in adult patients in Gabon
- Trial protocol pending to be included


<!-- \begin{table}[h!] -->
<!-- \centering -->
<!-- \begin{tabular}{|c|c|c|c|c|} -->
<!-- 	\hline -->
<!-- 	Disease & Endpoints & Analysis method & Sample size & Estimated responses \\ -->
<!-- 	\hline -->
<!-- 	Onchocerciasis &  &  &  &  \\ -->
<!-- 	\hline -->
<!-- 	Loiasis &  &  &  &  \\ -->
<!-- 	\hline -->
<!-- 	Mansonellosis &  &  &  &  \\ -->
<!-- 	\hline -->
<!-- 	Trichiuriasis &  &  &  &  \\ -->
<!-- 	\hline -->
<!-- \end{tabular} -->
<!-- \caption{} -->
<!-- \end{table} -->

\newpage

## Preliminary code to simulate data 

Next, we include a preliminary function to simulate basket trial data. In particular, the generated individual data includes the binary response and as covariates

- Substudy: we considered four substudies, which would correspond to each of the diseases in the basket trial
- Treatment: we considered two treatment arms and a control arm
- Patient's covariates: we included sex as covariate

The covariates treatment and sex are considered as fixed effects, while substudy is considered as random effect in this example. 

```{r, warning=F}
simulate_binary_data <- function(n, beta, b, ntreat=3, nsex=2, nstudies=4) {
  # add intercept and design matrix for fixed effects
  x <- matrix(rnorm(n), nrow = n, ncol = 1)
  if (ntreat==3) {
    treatments <- factor(sample(1:ntreat, n, replace = TRUE), 
                         levels = 1:ntreat, 
                         labels = c("Placebo", "Treatment 1", "Treatment 2"))
    X <- model.matrix(~ treatments - 1)
    x <- cbind(x, X)
  }
  if (nsex==2) {
    sex <- factor(sample(c("Male", "Female"), n, replace = TRUE))
    X <- model.matrix(~ sex - 1)
    x <- cbind(x, X)
  }
  # add random effects
  if (nstudies==4) {
    studies <- factor(sample(1:nstudies, n, replace = TRUE), 
                      levels = 1:nstudies, 
                      labels = c("S1", "S2", "S3", "S4"))
    Z <- model.matrix(~ studies - 1)
    rn = rnorm(n, mean = 0, sd = 1) 
    z <-  Z*rn
  } 
  
  # compute linear predictor as a function of covariates and coefficients 
  eta <- x %*% beta + z %*% b
  
  # convert linear predictor to probability using logistic function
  p <- plogis(eta)
  
  # simulate binary outcomes based on probability
  y <- rbinom(n, 1, p)
  
  # simulated data
  data <- data.frame(y, x, Z) 
  
  if (ntreat==3 && nsex==2 && nstudies==4) {
    data$treatments <- "T"
    data$treatments[data$treatmentsPlacebo==1] <- "Control"
    data$treatments[data$treatmentsTreatment.1==1] <- "Treat1"
    data$treatments[data$treatmentsTreatment.2==1] <- "Treat2"
    data$treatmentsPlacebo<-NULL
    data$treatmentsTreatment.1<-NULL
    data$treatmentsTreatment.2<-NULL 
    data$sex <- "S"
    data$sex[data$sexFemale==1] <- "F"
    data$sex[data$sexMale==1] <- "M"
    data$sexFemale<-NULL
    data$sexMale<-NULL 
    data$substudy <- "SS"
    data$substudy[data$studiesS1==1] <- "Study1"
    data$substudy[data$studiesS2==1] <- "Study2"
    data$substudy[data$studiesS3==1] <- "Study3"
    data$substudy[data$studiesS4==1] <- "Study4"
    data$studiesS1<-NULL
    data$studiesS2<-NULL
    data$studiesS3<-NULL
    data$studiesS4<-NULL 
    data$V1<-NULL
  }
  
  model <- glmer(y ~ treatments + sex + (1 | substudy), 
                 data = data, family = binomial) 
  
  return(list(data=data, model=model))
  
}
```

Example:
```{r} 
set.seed(2345)
ntreat=3; nsex=2; nstudies=4; n =4*1000 
beta <- c(1,0,0.5,1,0,0.2) 
b=c(.1,.2,.4,.5)
# 
sim_res <- simulate_binary_data(n=n, beta=beta, b=b, ntreat=3, nsex=2, nstudies=4)
head(sim_res$data,10)
```
Fitting the model
```{r} 
summary(sim_res$model)
ranef(sim_res$model)
```





