---
title: "Report -- eWHORM simulations"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Transformation of the data


**Step 1:** 

* Let's consider $x_0, x_1$ the variable denoting microfilaria at baseline and at twelve months  and $sd_0$
 the common standard deviation for a specific treatment. 

* $\mu_0 = mean(x_0)$, and $\mu_1 = (1 - r_0) \mu_0$
 
**Step 2:** 

* we consider the $log(x_1)$ and  $log(x_0)$ to calculate the difference: $\mu = log(x_1) - log(x_0)$ with

the following assumptions: 

* $\tilde{x_i} = log(x_i)$ with the estimated mean $m_i$ and the variance $\sigma_{\tilde{x_i}}^2$

* $(\tilde{x_0}, \tilde{x_1})$ following bivariate normal distribution with mean

$m = (m_0, m_1)$ and  variance-covariance matrix with elements $\sigma_{\tilde{x_0}}^2, \sigma_{\tilde{x_1}}^2, and \quad cov(\tilde{x_1}, \tilde{x_0}) =cov(\tilde{x_0}, \tilde{x_1}) = \rho \sigma_{\tilde{x_0}} \sigma_{\tilde{x_1}}$

* From the known formula we  obtained $m_i = log(\frac{\mu_i^2}{   \sqrt{\mu_i^2 + mean(sd_i)^2} })$,
            $\sigma_{\tilde{x_i}} = \sqrt{log(1 + \frac{sd_i^2}{\mu_i^2}  )})$,  $sd_0 = sd_i$, $\forall i$
            

* The difference of the log is following the normal distribution with mean = $m_1 - m_0$, var = $\sigma_{\tilde{x_0}}^2 + \sigma_{\tilde{x_1}}^2 - 2 \rho \sigma_{\tilde{x_1}} \sigma_{\tilde{x_0}} = \sigma_{\tilde{x_0}}^2$ ($\rho = \frac{1}{2}, \sigma_{\tilde{x_0}}^2 = \sigma_{\tilde{x_i}}^2)$


## Values for the simulations

baseline means, sd, correlations, ...

## Considered parameters for the design

Values

alpha1, alpha, power, sample sizes, ...

- code examples

Let's consider $mean(x_0) = 650, sd_0 = 575$


```{r pressure, echo=FALSE, results='asis'}

flog1 = function(mn = 650, sd = 575,  r0 = 0.10, r1 = 0.6, r2 = 0.8, r3 = 0.9, r = 0.5, rho = 0.5){
  
  
  #The mean after twelve months from the common baseline mn
  #########################################################
  
  mn0 = (1 - r0)*mn
  mn1 = (1 - r1)*mn
  mn2 = (1 - r2)*mn
  mn3 = (1 - r3)*mn
  
  
  
  sd0 = sd #we assumed the same sd for all the group
  
  moy0 = log(mn0^2/sqrt(mn^2 + sd0^2)); sde0 <- sqrt(log(1+(sd0^2/mn^2) ))
  moy1 = log(mn0^2/sqrt(mn0^2 + sd0^2)); sde0 <- sqrt(log(1+(sd0^2/mn1^2) ))
  moy2 = log(mn1^2/sqrt(mn1^2+sd0^2)); sde1 <- sde0 #sqrt(log(1+(sd1^2/mn1^2) ))
  moy3 =  log(mn2^2/sqrt(mn2^2+sd0^2)); sde2 <- sde0 #sqrt(log(1+ (sd2^2/mn2^2) ))
  moy4 =  log(mn3^2/sqrt(mn3^2+sd0^2)); sde3 <- sde0 
  
  #Variance of the diff of the logs
  #################################
  
  rho = 0.5; vardiff = sde1^2 + sde2^2 - 2* rho*sde1*sde2 
  
  # Then mean(log(xi) - log(x0))
  
  mu1 = moy1 - moy0
  mu2 = moy2 - moy0
  mu3 = moy3 - moy0
  mu4 = moy4 - moy0
  
  # The mu for the simulation: this is about the effect in log
  ############################################################
  
  #mu = log(c((r0)*x0, (r1)*x0, (r2)*x0, (r3)*x0))
  
  mu =  c(-log(1 - r0), -log(1 - r1), -log(1 - r2), -log(1 - r3))  #diff baseline and six months 
  
  #or 
  
  mu =  c(mu1, mu2, mu3, mu4)
  
  
  # variance-covariance matrix
  ############################
  
  a = rho*sde0^2 # covariance
  
  var = vardiff
  
  sg2 = matrix(c(var, a, a, var),nrow=2,byrow = T)
  
  return(list(mn, mu, var, sg2))
  
}
```


```{r pressure, echo=FALSE}



```
